\ProjectName{}: Data Imputation as a Query Optimization

### 2016-12-14 paper/algorithm.tex

$\text{Loss}(Q)$ is computed as follows.
For each sub-query, in addition to a histogram, we keep track of the estimated number of
null values in each column, $\text{MissingCount}(c, Q)$.
\begin{align*}
  \text{Loss}(Q) = \begin{cases}
    \sum_{c \in C} \text{MissingCount}(c, Q') & Q = \delta_C(Q') \\
    \frac{1}{\sqrt{|Q'|}} \sum_{c \in C} \text{MissingCount}(c, Q') & Q = \mu_C(Q') \\
    \text{Loss}(Q_1') + \text{Loss}(Q_2') & Q = Q_1' \bowtie_\psi Q_2' \\
    \text{Loss}(Q') & Q = \sigma_\phi(Q'), Q = \pi_C(Q') \\
  \end{cases}
\end{align*}
Dropping tuples which contain a null value incurs a loss of 1 for each null field dropped.
Imputing null fields incurs a loss penalty $p \in (0, 1]$ which decreases as the number of
tuples available to train the imputation algorithm increases.  Intuitively, imputation
accuracy \textit{increases} when more complete attributes and complete tuples are available;
correspondingly, the information loss \textit{decreases} with more complete values to work
with. The inverse square loss can be seen as a first approximation to the learning guarantees of
regression trees. \todo{Micah: Probably need to look into this more carefully since I don't
know much about the learning guarantees of regression trees!} Since we conceive that
alternate imputation strategies could be used, the analyst would provide a corresponding
loss with a reasonable parameterization.

\todo{Micah: Notation not super clear to me. I'm guessing $|Q'|$ is the number of tuples total, but
$Q'$ is a query plan, and I would interpret it's cardinality as perhaps the height of the
tree or number of elements in the tree, not in terms of underlying relation. Similarly, I'm
not sure how to define $\text{MissingCount}(\cdot)$ above.}

>>>>

$\text{Loss}(Q)$ is computed as follows.
For each sub-query, using the histogram information, we keep track of the estimated number of
null values in each column, $\text{MissingCount}(c, Q)$.
\begin{align*}
  \text{Loss}(Q) = \begin{cases}
    \sum_{c \in C} \text{MissingCount}(c, Q') & Q = \delta_C(Q') \\
    \frac{1}{\sqrt{|Q'|}} \sum_{c \in C} \text{MissingCount}(c, Q') & Q = \mu_C(Q') \\
    \text{Loss}(Q_1') + \text{Loss}(Q_2') & Q = Q_1' \bowtie_\psi Q_2' \\
    \text{Loss}(Q') & Q = \sigma_\phi(Q'), Q = \pi_C(Q') \\
  \end{cases}
\end{align*}
Dropping tuples which contain a null value incurs a loss of 1 for each null field dropped.
Imputing null fields incurs a loss penalty $p \in (0, 1]$ which decreases as the number of
tuples available to train the imputation algorithm increases.  Intuitively, imputation
accuracy \textit{increases} when more complete attributes and complete tuples are available;
correspondingly, the information loss \textit{decreases} with more complete values to work
with. The inverse square loss can be seen as a heuristic first approximation to the learning guarantees of
regression trees. Since we conceive that
alternate imputation strategies could be used, the analyst would provide a corresponding
loss with a reasonable parameterization.
